NLTK: The Natural Language Toolkit

### 2004-04-30 technical/acl2004/acl04.tex
- Minor edits (noded by [el] comments) - Made code samples \small
svn/trunk@1780

\textbf{Assignments.}
NLTK supports assignments of varying difficulty
and scope.  In the simplest assignments, students experiment with
existing components for a wide variety of NLP tasks.  As students
become more familiar with the toolkit, they can be asked to modify
existing components, or to create complete systems out of existing
components.

@@

We chose Python because it has a shallow learning curve, its syntax
and semantics are transparent, and it has good string-handling
functionality.  As an interpreted language, Python facilitates
interactive exploration.  As an object-oriented language, data and
methods can be encapsulated easily.  Python's extensive standard
library is useful makes numerical and graphical programming as easy as
it could be, and the recently added generator syntax makes it easy to
create interactive implementations of algorithms
\cite{Loper04,Rossum03intro,Rossum03ref}.

\begin{figure}[bth]
\epsfig{file=../pycon2004/chart.eps, width=\linewidth}
\caption{Interactive Chart Parsing Demonstration}
\label{fig:chart}
\end{figure}

@@

NLTK is implemented as a large collection of minimally interdependent
modules, organized into a shallow package hierarchy.  A set of core
modules defines basic data types that are used throughout the toolkit.  
The remaining modules are \emph{task modules}, each devoted to an
individual natural language processing task.  For example, the
\texttt{nltk.parser} package encompasses to the task of
\emph{parsing}, or deriving the syntactic structure of a sentence; and
the \texttt{nltk.tokenizer} module is devoted to the task of
\emph{tokenizing}, or dividing a text into its constituent parts.

@@

To maximize interoperability between different modules, we use a
single class to encode information about natural language texts -- the
\texttt{Token} class.  Each \texttt{Token} instance represents a
single unit of text, such as a word, sentence, or document; and is
defined by (partial) mapping from property names to values.  For
example, the \texttt{TEXT} property is used to encode a token's text
content:

\begin{alltt}
\textbf{>>> from nltk.token import *}
\textbf{>>> Token(TEXT="Hello World!")}
<Hello World!>
\end{alltt}
%
The \texttt{TAG} property is used to encode a token's part-of-speech
tag:

\begin{alltt}
\textbf{>>> Token(TEXT="python", TAG="NN")}
<python/NN>
\end{alltt}
%
The \texttt{SUBTOKENS} property is used to store a tokenized text:

\begin{alltt}
\textbf{>>> from nltk.tokenizer import *}
\textbf{>>> token = Token(TEXT="Hello World!")}
\textbf{>>> WSTokenizer().tokenize(token)} \textit{\# whitespace tokenizer}
\textbf{>>> print token['SUBTOKENS'])}
[<Hello>, <World!>]
\end{alltt}
%
In a similar fashion, other language processing tasks such as
word-sense disambiguation, chunking and parsing all add properties to
the \texttt{Token} data structure.

@@

\begin{alltt}
\textbf{>>> from nltk.corpus import brown}
\textbf{>>> brown.items()}
['ca01', 'ca02', 'ca03', 'ca04', 'ca05', 'ca06',
'ca07', 'ca08', 'ca09', 'ca10', 'ca11', 'ca12', ... ]
\textbf{>>> brown.tokenize('ca01')}
<[<The/at>, <Fulton/np-tl>, <County/nn-tl>,
<Grand/jj-tl>, <Jury/nn-tl>, <said/vbd>, <Friday/nr>,
<an/at>, <investigation/nn>, <of/in>, <Atlanta's/np$>,
<recent/jj>, <primary/nn>, <election/nn>,
<produced/vbd>, <``/``>, <no/at>, <evidence/nn>,
<''/''>, <that/cs>, <any/dti>, <irregularities/nns>,
<took/vbd>, <place/nn>, <./.>, ...]>
\end{alltt}
%
A selection of 5\% of the Penn Treebank corpus is included with
NLTK, and it can be accessed as follows:

\begin{alltt}
\textbf{>>> from nltk.corpus import treebank}
\textbf{>>> treebank.items()}
['wsj_0001.prd', 'wsj_0002.prd', 'wsj_0003.prd', ...]
\textbf{>>> treebank.tokenize('parsed/wsj_0001.prd')}
[ (S: (NP-SBJ: (NP: <Pierre> <Vinken>) ...)
        (VP: will ...)),
  (S: (NP-SBJ: <Mr.> <Vinken>)
        (VP: <is> ...)) ]
\end{alltt}
%
[WHY CAN'T I PP() OR DRAW() THE RESULTING OBJECT?]

@@

Three different types of documentation are available.  Tutorials
explain how to use the toolkit, with detailed worked examples.  The
API documentation describes every module, interface, class, method,
function, and variable in the toolkit.  Technical reports explain and
justify the toolkit's design and implementation.  All are available
from \url{http://nltk.sourceforge.net/docs.html}.

@@

NLTK is available from \url{nltk.sourceforge.net}, and is packaged for
easy installation under Unix, Mac OS X and Windows.  The full
distribution consists of four packages: the Python source code
(\texttt{nltk}); the corpora (\texttt{nltk-data}); the documentation
(\texttt{nltk-docs}); and third-party contributions
(\texttt{nltk-contrib}).  Before installing NLTK, it is necessary to
install Python version 2.3 or later, available from
\url{www.python.org}.  Full installation instructions and a quick
start guide are available from the NLTK homepage.

As soon as NLTK is installed, users can run the demonstrations.  On
Windows, the demonstrations can be run by double-clicking on their
Python source files.  Alternatively, from the Python interpreter, this
can be done as follows:

\begin{alltt}
\textbf{>>> import nltk.draw.rdparser}
\textbf{>>> nltk.draw.rdparser.demo()}
\textbf{>>> nltk.draw.srparser.demo()}
\textbf{>>> nltk.draw.chart.demo()}
\end{alltt}

@@

NLTK is an open source project, and we welcome any contributions.
There are several ways to contribute: users can report bugs, suggest
features, or contribute patches on Sourceforge; users can participate
in discussions on the NLTK-Devel mailing list or in the NLTK public
forums; and users can submit their own NLTK-based projects for
inclusion in the nltk\_contrib directory.  New code modules that are
relevant, substantial, original and well-documented will be considered
for inclusion in NLTK proper.  Potential contributors should note that
all source code is distributed under the Gnu General Public License,
and all documentation is distributed under a Creative Commons
non-commercial license.  Thus they can be confident that their
work will remain freely available to all.  Further information about
contributing to NLTK is available at \url{http://nltk.sourceforge.net/contrib.html}.

@@

NLTK is a broad-coverage toolkit that provides a simple, extensible,
uniform framework for assignments, demonstrations and projects.  It is
thoroughly documented, easy to learn, and simple to use.  Readers who
would like to receive occasional announcements about NLTK are
encouraged to sign up for the low-volume, moderated mailing list
\textit{NLTK-Announce}.

>>>>

\textbf{Assignments.}
NLTK supports assignments of varying difficulty
and scope.  In the simplest assignments, students experiment with
existing components to perform a wide variety of NLP tasks.  As students
become more familiar with the toolkit, they can be asked to modify
existing components, or to create complete systems out of existing
components.

@@

We chose Python because it has a shallow learning curve, its syntax
and semantics are transparent, and it has good string-handling
functionality.  As an interpreted language, Python facilitates
interactive exploration.  As an object-oriented language, data and
methods can be encapsulated easily.  Python comes with an extensive
standard library, including tools for graphical programming and
intensive numerical processing.  And the recently added generator
syntax makes it easy to create interactive implementations of
algorithms \cite{Loper04,Rossum03intro,Rossum03ref}.

\begin{figure}[bth]
\epsfig{file=../pycon2004/chart.eps, width=\linewidth}
% [el] why the ``../pycon2004'' in the file name?
\caption{Interactive Chart Parsing Demonstration}
\label{fig:chart}
\end{figure}

@@

% [el] I got rid of the package/module distinction here, because I 
% think it'll just be confusing to this audience.  So let's just
% call everything a module (and say we have a module hierarchy).
NLTK is implemented as a large collection of minimally interdependent
modules, organized into a shallow hierarchy.  A set of core
modules defines basic data types that are used throughout the toolkit.
The remaining modules are \emph{task modules}, each devoted to an
individual natural language processing task.  For example, the
\texttt{nltk.parser} module encompasses to the task of
\emph{parsing}, or deriving the syntactic structure of a sentence; and
the \texttt{nltk.tokenizer} module is devoted to the task of
\emph{tokenizing}, or dividing a text into its constituent parts.

@@

To maximize interoperability between different modules, we use a
single class to encode information about natural language texts -- the
\texttt{Token} class.  Each \texttt{Token} instance represents a
single unit of text, such as a word, sentence, or document; and is
defined by a (partial) mapping from property names to values.  For
example, the \texttt{TEXT} property is used to encode a token's text
content:

\begin{alltt}\small
\textbf{>>> from nltk.token import *}
\textbf{>>> Token(TEXT="Hello World!")}
<Hello World!>
\end{alltt}
%
The \texttt{TAG} property is used to encode a token's part-of-speech
tag:

\begin{alltt}\small
\textbf{>>> Token(TEXT="python", TAG="NN")}
<python/NN>
\end{alltt}
%
The \texttt{SUBTOKENS} property is used to store a tokenized text:

% [el] I'm assuming that 1.4 will be released by ACL, so use
% WhitespaceTokenizer instead of WSTokenizer.
\begin{alltt}\small
\textbf{>>> from nltk.tokenizer import *}
\textbf{>>> token = Token(TEXT="Hello World!")}
\textbf{>>> WhitespaceTokenizer().tokenize(token)}
\textbf{>>> print token['SUBTOKENS'])}
[<Hello>, <World!>]
\end{alltt}
%
In a similar fashion, other language processing tasks such as
word-sense disambiguation, chunking and parsing all add properties to
the \texttt{Token} data structure.

@@

% [el] Changed to show off CorpusReader.groups().
\begin{alltt}\small
\textbf{>>> from nltk.corpus import brown}
\textbf{>>> brown.groups()}
['skill and hobbies', 'popular lore', 
'humor', 'fiction: mystery', ...]
\textbf{>>> brown.items('humor')}
('cr01', 'cr02', 'cr03', 'cr04', 'cr05',
'cr06', 'cr07', 'cr08', 'cr09')
\textbf{>>> brown.tokenize('cr01')}
<[<It/pps>, <was/bedz>, <among/in>,
<these/dts>, <that/cs>, <Hinkle/np>,
<identified/vbd>, <a/at>, ...]>
\end{alltt}
%
A selection of 5\% of the Penn Treebank corpus is included with
NLTK, and it can be accessed as follows:

% [el] Changed to show groups() and TreeToken.pp()
% [el] This example wraps over a page at a weird place.  When
% we've made all the other changes, maybe manually insert a
% page break at a better location?
\begin{alltt}\small
\textbf{>>> from nltk.corpus import treebank}
\textbf{>>> treebank.groups()}
('raw', 'tagged', 'parsed', 'merged')
\textbf{>>> treebank.items('parsed')}
['wsj_0001.prd', 'wsj_0002.prd', ...]
\textbf{>>> item = 'parsed/wsj_0001.prd'}
\textbf{>>> sentences = treebank.tokenize(item)}
\textbf{>>> for sent in sentences['SUBTOKENS']:}
\textbf{...     print sent.pp()} \emph{# pretty-print}
(S:
  (NP-SBJ:
    (NP: <Pierre> <Vinken>)
    ...
\end{alltt}
%

@@

% [el] Changed the url here and elsewhere from nltk.sourceforge.net
% to nltk.sf.net, because it fits better on small columns (and because 
% it brought the length back down to 4 pages :) ).
Three different types of documentation are available.  Tutorials
explain how to use the toolkit, with detailed worked examples.  The
API documentation describes every module, interface, class, method,
function, and variable in the toolkit.  Technical reports explain and
justify the toolkit's design and implementation.  All are available
from \url{http://nltk.sf.net/docs.html}.

@@

NLTK is available from \url{nltk.sf.net}, and is packaged for
easy installation under Unix, Mac OS X and Windows.  The full
distribution consists of four packages: the Python source code
(\texttt{nltk}); the corpora (\texttt{nltk-data}); the documentation
(\texttt{nltk-docs}); and third-party contributions
(\texttt{nltk-contrib}).  Before installing NLTK, it is necessary to
install Python version 2.3 or later, available from
\url{www.python.org}.  Full installation instructions and a quick
start guide are available from the NLTK homepage.

% [el] We really should make a better interface for running demos.
% But I don't think we can have that figured out by the deadline
% for this paper.
As soon as NLTK is installed, users can run the demonstrations.  On
Windows, the demonstrations can be run by double-clicking on their
Python source files.  Alternatively, from the Python interpreter, this
can be done as follows:

\begin{alltt} \small
\textbf{>>> import nltk.draw.rdparser}
\textbf{>>> nltk.draw.rdparser.demo()}
\textbf{>>> nltk.draw.srparser.demo()}
\textbf{>>> nltk.draw.chart.demo()}
\end{alltt}

@@

NLTK is an open source project, and we welcome any contributions.
There are several ways to contribute: users can report bugs, suggest
features, or contribute patches on Sourceforge; users can participate
% [el] Add a \cite with a url, or a url inline?  The url is:
% http://lists.sourceforge.net/lists/listinfo/nltk-devel
in discussions on the \textit{NLTK-Devel} mailing list or in the NLTK public
forums; and users can submit their own NLTK-based projects for
inclusion in the nltk\_contrib directory.  New code modules that are
relevant, substantial, original and well-documented will be considered
for inclusion in NLTK proper.  Potential contributors should note that
% [el] Should it be GNU instead of Gnu?
all source code is distributed under the Gnu General Public License,
and all documentation is distributed under a Creative Commons
non-commercial license.  Thus they can be confident that their
work will remain freely available to all.  Further information about
contributing to NLTK is available at \url{http://nltk.sf.net/contrib.html}.

@@

NLTK is a broad-coverage toolkit that provides a simple, extensible,
uniform framework for assignments, demonstrations and projects.  It is
thoroughly documented, easy to learn, and simple to use.  Readers who
would like to receive occasional announcements about NLTK are
encouraged to sign up for the low-volume, moderated mailing list
% [el] Add a \cite with a url, or a url inline?  The url is:
% http://lists.sourceforge.net/lists/listinfo/nltk-announce
\textit{NLTK-Announce}.
